---
title: "world_happiness report 2024"
author: "Fatma TNANI"
format: html
---

## Loading Libraries and Data
First, let's load the necessary libraries and import the dataset
```{r}
#| message: false
library(tidyverse)
library(readr)

# reading the csv file
happiness_data <- read.csv("World-happiness-report-updated_2024.csv")

# Dis^play the first few rows and structure of the data
head(happiness_data)
str(happiness_data)
```
## Examining the Data

Let's check if our data adheres to tidy data principles:

1-Each variable forms a column

2-Each observation forms a row

3-Each type of observational unit forms a table

```{r}
# check for any missing values
colSums(is.na(happiness_data))

# check the number of unique countries and years
n_distinct(happiness_data$'Country name')
n_distinct(happiness_data$year)
```

## Tidying the Data

Our data seems to be mostly tidy, but we can make a few improvements:

Rename columns to remove spaces and make them more R-friendly

Separate the "Dystopia + residual" column into two columns

Create separate columns for positive and negative affects if they're combined

```{r}
tidy_happiness <- happiness_data |> 
  #Ren  
```

